% §6 Related Work
% Structure: MemSkill §2 sentence-level skeleton
% Prose quality: GEPA-level (causal chains, specific mechanisms, dense)
% Target: ~314 words ± 15%

\section{Related Work}

\noindent\textbf{Memory for LLM Agents.}
Prior work on agent memory builds external stores from interaction histories~\citep{zhang2024survey_memory}: typical pipelines extract salient information, retrieve relevant entries for a new query, and update via consolidation or pruning~\citep{park2023generative,zhong2024memorybank,wang2024augmenting}.
Recent systems learn richer policies; MemSkill~\citep{memskill} selects natural-language skills via an RL controller, and AWM~\citep{wang2024awm} induces reusable workflows from successful trajectories.
Concurrent work on self-evolving memory stays within closed design spaces, e.g., MemEvolve~\citep{memevolve} searches over four predefined modules (encode, store, retrieve, manage).
All commit to a fixed representation at design time and cannot express task-specific computational logic such as relational schemas or stateful aggregations.
By contrast, we search over executable Python programs that implement arbitrary data structures and retrieval logic, subsuming these closed spaces.

\noindent\textbf{Self-Evolving Agents.}~\looseness=-1
Self-evolving LLM agents improve from interaction experience without gradient-based updates~\citep{gao2025survey_selfevolving}.
ExpeL~\citep{zhao2024expel} distills trajectories into editable natural-language insights, FLEX~\citep{cai2025flex} builds an experience library via continual reflection, and a complementary line automates agent design through meta-optimization (ADAS~\citep{hu2024adas}), skill synthesis (Voyager~\citep{wang2023voyager}), and verbal self-critique (Reflexion~\citep{shinn2023reflexion}).
These systems evolve text-level artifacts that the LLM must re-interpret at runtime; \sysname{} instead evolves deterministic code, reducing interpretation variance and enabling explicit data structures and retrieval logic in each memory program.

\noindent\textbf{Program Search with Language Models. }
Using LLMs as mutation operators has yielded solutions surpassing human heuristics in combinatorial optimization~\citep{romera2024funsearch}, neural architecture search~\citep{chen2023evoprompting}, and algorithmic discovery~\citep{deepmind2025alphaevolve}.
Closest to our work, GEPA~\citep{gepa} mutates every prompt in a compound AI system via reflective natural-language updates with Pareto-front candidate selection.
\sysname{} applies the same paradigm to executable memory programs; because generated code can fail to compile or violate runtime constraints, we add a compile-fix loop and runtime violation recovery that are unnecessary in prompt evolution (Section~\ref{sec:reflection}).
