\documentclass{article}

% ============================================================
% TERMINOLOGY & NOTATION CONVENTIONS
% ============================================================
%
% System name:  \sysname  (Mem★, rendered as \textsc{Mem}$\star$)
%
% --- Terms ---
%   memory harness   - high-level framing term for task-facing memory machinery.
%                      Use only in the title, abstract, introduction, and conclusion.
%   memory program   - formal object optimized by \sysname{}; the evolvable Python module ($P$).
%                      Use for definitions, equations, method, results, transfer, and ablations.
%   memory-program search
%                    - formal problem: searching over executable memory programs.
%   reflective code evolution
%                    - method used by \sysname{} to solve memory-program search.
%   knowledge base   - runtime data store populated by a program ($\mathcal{K}$)
%
% --- Program components (four; used as \textbf{} paragraph headers) ---
%   Schema      — KnowledgeItem + Query dataclass fields (what the agent parses)
%   Instruction — four string constants guiding the task agent
%   Logic       — write() + read() implementation (storage & retrieval code)
%   Toolkit     — SQLite, ChromaDB, LLM primitives available to programs
%
% --- Global symbols (keep to ~7) ---
%   $P$              memory program (being evolved)
%   $\mathcal{A}$    task agent (frozen weights)
%   $\mathcal{K}$    knowledge base instance
%   $J(P)$           fitness score of program $P$
%   $D_\text{e}$     episodes used for write()
%   $D_\text{v}$     validation queries for fitness scoring
%   $D_\text{t}$     held-out test queries (final report only)
%   $\mathcal{P}$    program pool
%
% --- Local-only symbols (define where used, don't treat as global) ---
%   $\mu$            task metric (only in $J(P)$ definition)
%   $\tau$           softmax temperature (only in selection eq.)
%   $B$              evaluation budget (only in method)
%   $\mathcal{R}$    reflection context (only in mutation eq.)
%
% --- Style rules ---
%   .read() / .write()  — always \texttt{read} / \texttt{write} (never .retrieve())
%   Data subscripts      — always \text{}, e.g. $D_\text{val}$ (not $D_s$ or $D_{val}$)
%   System name          — always \sysname (never plain text "MemStar")
% ============================================================

\newif\ifmstarpreprint
\mstarpreprintfalse
% \mstarpreprinttrue

\PassOptionsToPackage{numbers, compress}{natbib}
\ifmstarpreprint
  \usepackage[preprint]{neurips_2026}
\else
  \usepackage{neurips_2026}
\fi
\raggedbottom

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{CJKutf8}
\newcommand{\cm}[1]{\begin{CJK}{UTF8}{gbsn}#1\end{CJK}}
\usepackage{hyperref}
\hypersetup{hidelinks}
\usepackage{url}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage[table]{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{placeins}
\usepackage{listings}
\usepackage{pifont}
\usepackage{fontawesome5}
\usepackage{tikz}
\usetikzlibrary{calc,shapes.geometric}
\ifmstarpreprint
  \usepackage[most]{tcolorbox}
  \usepackage{varwidth}
\fi
\usepackage{enumitem}
\setlist{leftmargin=*}

\ifmstarpreprint
% Callout-box abstract
\definecolor{abstractbg}{gray}{0.97}
\renewenvironment{abstract}{%
  \begin{tcolorbox}[
    enhanced,
    breakable,
    colback=abstractbg,
    colframe=black!15,
    boxrule=0.4pt,
    arc=4pt,
    left=6pt,
    right=6pt,
    bottom=6pt,
    top=6pt,
    shadow={1pt}{-1pt}{0pt}{black!8},
  ]
  \small\noindent\tcbox[on line, colback=black!8, colframe=black!8, boxrule=0pt, arc=2pt, left=2pt, right=2pt, top=1pt, bottom=1pt]{\textbf{\normalsize Abstract}}\enspace
}{%
  \end{tcolorbox}
}
\fi

\definecolor{codebg}{HTML}{F7FAFC}
\definecolor{codeframe}{HTML}{93C5FD}
\definecolor{codeaccent}{HTML}{2563EB}
\definecolor{codecomment}{HTML}{64748B}
\definecolor{codestring}{HTML}{047857}
\definecolor{codekeyword}{HTML}{1D4ED8}

\captionsetup[lstlisting]{
  font=small,
  labelfont={bf,color=codeaccent},
  textfont=bf,
  singlelinecheck=false,
  skip=4pt,
}

% Code listing style for evolved program callouts
\lstdefinestyle{evolvedcode}{
  language=Python,
  basicstyle=\ttfamily\scriptsize\linespread{1.05}\selectfont,
  keywordstyle=\bfseries\color{codekeyword},
  commentstyle=\itshape\color{codecomment},
  stringstyle=\color{codestring},
  numbers=none,
  frame=single,
  frameround=tttt,
  framesep=5pt,
  framerule=0.45pt,
  rulecolor=\color{codeframe},
  backgroundcolor=\color{codebg},
  breaklines=true,
  showstringspaces=false,
  columns=flexible,
  keepspaces=true,
  aboveskip=0.75em,
  belowskip=0.6em,
  xleftmargin=0.6em,
  xrightmargin=0.6em,
  framexleftmargin=0.55em,
  framexrightmargin=0.55em,
  framextopmargin=0.25em,
  framexbottommargin=0.25em,
  captionpos=t,
  escapeinside={(*@}{@*)},
  morekeywords={self},
}

% Scaffold marker: writing agents replace \scaffold{} blocks with real text
\newcommand{\scaffold}[1]{\begingroup\color{red}\textbf{[SCAFFOLD]} #1\endgroup}

% System name
\newcommand{\sysname}{\textsc{Mstar}}
% Second-best result highlighting
\newcommand{\second}[1]{\textcolor{green!50!black}{\textbf{#1}}}

\title{M\textsuperscript{$\bigstar$}: Every Task Deserves Its Own Memory Harness}

\author{
  \textbf{Wenbo Pan}$^{1}$\thanks{Correspondence to: \texttt{wenbo.pan@my.cityu.edu.hk}} \quad
  \textbf{Shujie Liu}$^{2}$ \quad
  \textbf{Xiangyang Zhou}$^{2}$ \quad
  \textbf{Shiwei Zhang}$^{2}$ \quad
  \textbf{Wanlu Shi}$^{2}$ \\[0.3em]
  \textbf{Mirror Xu}$^{2}$ \quad
  \textbf{Xiaohua Jia}$^{1}$ \\[0.5em]
  $^{1}$City University of Hong Kong \qquad
  $^{2}$Microsoft
}

\begin{document}

\maketitle

% === Visual Abstract (Hero Figure): Evolved memory programs across tasks ===
\input{figures/overview/evolved_programs}

\begin{abstract}
Large Language Model (LLM) agents rely on a memory harness to write, organize, retrieve, and use past experience.
A harness that works for one task often fails on another, because different tasks like conversations, embodied planning, or expert reasoning require different storage and retrieval behavior.
To address this limitation, we introduce \sysname{}, a method that automatically discovers task-optimized memory programs through reflective code evolution.
Specifically, \sysname{} represents a memory harness as an executable Python program that defines the \textit{Schema}, \textit{Logic}, and \textit{Instruction} components, which are optimized jointly.
We evaluate \sysname{} on four tasks covering conversation, embodied planning, and specialized reasoning.
Our results demonstrate that \sysname{} improves performance over existing static memory harnesses across all evaluated tasks.
Furthermore, the evolved memory programs exhibit structurally distinct processing mechanisms for each evaluated domain.
These findings suggest that every task benefits from its own memory harness, and that memory-program search provides a concrete way to discover it.
\end{abstract}

% === §1 Introduction ===
\section{Introduction}

Agentic systems powered by large language models (LLMs) need a memory harness to reuse past experience across diverse tasks~\citep{zhang2024survey_memory,park2023generative,zhong2024memorybank,wang2024augmenting}.
We use \textit{memory harness} to refer to the components that store, organize, retrieve, and use this experience.
As different task domains require knowledge in different structures, they call for different memory harnesses.
For example, conversational agents need semantic maps of people, events, and preferences to remember user-specific facts (Figure~\ref{fig:system-overview}, top right)~\citep{locomo}, while embodied AI agents need reusable patterns from past trajectories to choose actions in new environments (Figure~\ref{fig:system-overview}, bottom left)~\citep{alfworld}, and specialized reasoning domains may need structured databases (Figure~\ref{fig:system-overview}, top left)~\citep{prbench}.
These requirements encode different assumptions about what to store, how to organize it, and when to retrieve it.
A harness built around one set of assumptions can therefore transfer poorly to another~\citep{structmemeval}.
This raises a practical question: \textit{How can we find the most suitable memory harness for a given task?}

To answer this question, we propose \sysname{}, an evolutionary method that automatically discovers task-optimized memory harnesses for LLM agents.
Specifically, \sysname{} represents a memory harness as a \textit{memory program}: a Python module with a fixed interface consisting of three components, \textit{Schema}, \textit{Logic}, and \textit{Instruction}.
The \textit{Schema} defines the data formats for storage and retrieval, the \textit{Logic} implements the \texttt{read} and \texttt{write} operations, and the \textit{Instruction} guides the agent's interaction with the knowledge base.
These components are jointly optimized through reflective code evolution.
In comparison, although prior works have also searched over and improved agent memory harnesses, their search is typically confined to a narrow, designer-specified space~\citep{zhao2024expel,cai2025flex,memskill,memevolve,gepa}.
\sysname{} searches over the memory harness itself with two mechanisms: a reflector agent that debugs memory program failures and generates targeted improvements, and a population-based program search strategy that balances exploration and exploitation while enforcing constraints on program quality.


We apply \sysname{} to four tasks covering conversation, embodied planning, and specialized reasoning, evaluating its performance against nine competitive baselines across three memory paradigms~\citep{locomo,alfworld,healthbench,prbench}.
Our results indicate that \sysname{} achieves the best overall score and performs robustly on every task; in contrast, the baselines perform well only on specific tasks.
Furthermore, we observe that the optimal memory programs for different tasks are distinct in both structure and procedure.
By exploring a broad memory-program search space, \sysname{} evaluates diverse memory programs to identify the task-optimized memory program.
Together, these results support the view that every task benefits from its own memory harness, discovered as an executable memory program rather than chosen as a fixed design.

In summary, our main contributions are as follows:
\begin{enumerate}
    \item We formulate task-optimized memory harness discovery as \textit{memory-program search}, where an executable Python program jointly specifies Schema, Logic, and Instruction.
    \item We propose \sysname{}, an evolutionary method that optimizes memory harnesses from target-task feedback through reflective code evolution and achieves the strongest overall performance across four tasks compared to competitive generalist memory harnesses.
    \item We perform in-depth analysis over the memory harness space and find that task-optimized memory harnesses are distinct in both structure and procedure across different tasks: conversation, embodied planning, and expert reasoning tasks all require different optimal memory harnesses.
\end{enumerate}

% === §2 Problem Setting ===
\input{figures/method_overview}
\input{sections/problem}

% === §3 Method ===
\input{sections/method}

% === §4 Experimental Setup ===
\input{sections/setup}

% === §5 Results ===
\input{sections/results}

% === §6 Discussion ===
\input{sections/discussion}

% === §7 Related Work ===
\input{sections/related}

% === §8 Conclusion ===
\section{Conclusion}

We introduced \sysname{}, a method that discovers task-optimized memory programs for large language model agents through reflective code evolution.
By representing a memory harness as executable Python code with Schema, Logic, and Instruction, \sysname{} turns memory design into a memory-program search problem.
Across four benchmarks, the evolved programs outperform strong memory baselines and develop distinct storage, retrieval, and action workflows for different domains.
These results support the view that agent memory should be optimized for the target task, and that memory-program search provides a practical path toward this goal.
Future work will study more sample-efficient search strategies and extend this framework to broader classes of agent tasks.

\bibliographystyle{plainnat}
\bibliography{references}

\input{sections/appendix}

\clearpage
\input{checklist}

\end{document}
